---
title: "Strings and Factors"
author: "Elaine Yanxi Chen"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## String vectors

`str_detect` does not look for a complete match.

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_replace(string_vec, "jeff", "Jeff")

str_replace(string_vec, "m", "M")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```

When you would need to use regular expression?

There is the phrase "i think" in all four instances.
"^" is the beginning of things
"$" indicate the end of the line

Use when variable names have specific prefixes or suffixes

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "[Bb]ush")
```

Upper case or lower case b both would satisfy. Anything in this "[]" would be identified. 

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

Number followed by a letter or more than one letter.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Detect anything at all between 7 and 11.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[[0-9]")
```

How to search for an actual dot? search for a special character with "\\[" while "\\[[0-9]" search for a bracket and a number. 

## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

as.numeric(factor_vec)

factor_vec = fct_relevel(factor_vec, "male")
as.numeric(factor_vec)

```


## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

tidy up the NSDUH data ...

```{r}
marj_df = 
  table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    percent = str_replace(percent, "[a-b]$", ""),
    percent = as.numeric(percent),
    year = str_replace(year, "\\)", "")
  ) %>% 
  filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West", "District of Columbia"))
  )
```



```{r}
marj_df %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


## Restaurant inspections ...

```{r}
data("rest_inspec")
```

```{r}
rest_inspec
```

```{r}
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = grade, 
    values_from = n_obs
  )
```

```{r}
rest_inspec =
  rest_inspec %>% 
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))
  
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = grade, 
    values_from = n_obs
  )

```


Let's find pizza places ...

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  ggplot(aes(x = boro)) + 
  geom_bar()
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro), 
    boro = fct_recode(boro, "The City" = "Manhattan")) %>% 
  ggplot(aes(x = boro)) + 
  geom_bar()
```




